### 2023-02-15 talk/talk.tex
more on covid example

@@

  \begin{itemize}
  \item Metapopulation data were used to infer the fraction of asymptomatic cases and their contagiousness, using an iterated EnKF \\
    (Li et al, {\it Science}, May 2020).
   \item
    The time interval predates Wuhan travel restrictions.
    %
%     load("data/incidence.rda")
%    li20dat <- incidence[1:14,-1]
%     tot <- sort(apply(li20dat,2,sum))
%     sum(tot)         
   \item Total cases 801, Wuhan 454, Chongqing 27, Beijing 26.
   \item 259 cities with 0 cases, 99 with 1--5 cases, 17 with $>5$ cases.
\end{itemize}

@@

\nocite{asfaw23arxiv,bjornstad01,breto19,grenfell04,he10,ionides21,ionides22,lee20,li20,ng02,ning23-ibpf,park20,rebeschini15,wheeler23}

>>>>

\vspace{-4mm}

@@

  \begin{itemize}
  \item Metapopulation data were used to infer the fraction of asymptomatic cases and their contagiousness, using an iterated EnKF \\
    (Li et al, {\it Science}, May 2020).
   \item
    The time interval predates Wuhan travel restrictions.
%% see README for R code
   \item Total cases 801, Wuhan 454, Chongqing 27, Beijing 26.
   \item{} 259 cities with 0 cases, 99 with 1--5 cases, 17 with $>5$ cases.
   \item{} 0 cases reported for all cities during Jan 10--15.

\begin{frame}{Top: data. Middle: IBPF fit. Bottom: IEnKF fit} 
\includegraphics[width=12cm]{covid/3-panel.pdf}
\end{frame}

\begin{frame}{Technical issues with EnKF}

  \begin{itemize}
  \item EnKF is based on a continuous Gaussian approximation.
  \item Log-likelihoods with respect to counting measure are never positive, and cannot properly be compared with log-likelihoods corresponding to continuous densities.
    \item For data with many zeros, the unbounded EnKF likelihood can substantially bias the MLE.
    \item Adapting EnKF for count data is non-trivial (Katzfuss et al, {\it JASA}, 2019).
    \item Li et al avoided the issue by using a ``heuristic observation error variance'' within EnKF. The corresponding ``log-likelihood'' is not the log-likelihood of any model. In particular, it cannot validly be compared againt simple statistical benchmarks.
  \item Data and models will never be perfect. We want to reveal, rather than conceal, limitations.
  \end{itemize}

@@

\nocite{asfaw23arxiv,bjornstad01,breto19,grenfell04,he10,katzfuss19,ionides21,ionides22,lee20,li20,ng02,ning23-ibpf,park20,rebeschini15,wheeler23}
